rbcL ASV clustering

### 2019-11-05 code/rbcl_asv_clustering.Rmd
minor changes, re-named var

otus_asv_name <- otus %>% 
  left_join(otu_names, by="otu") %>% 
  select(asv_id, otu_name)
```

```{r echo=F, fig.width=8, fig.height=6}
tsne_edited_otus %>% 
  select(plot_taxa, otu) %>% 
  distinct() %>% 
  arrange(otu)

# make a key for downstream use (write to file at end of script)
asv_otu_key <- tsne_otu_names %>% 
  select(asv_id, otu, otu_name)

ggplot(tsne_otu_names, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), alpha=0.3, show.legend = FALSE) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_otu_names, is.na(plot_taxa)), 
             color = "grey",alpha = 0.8, size = 3) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_otu_names, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order)), alpha = 0.8, size = 3) +
  labs(title = "Points colored by both genus and species", color="ASV taxonomic assignment") +
  theme_few() 
```

@@

tsne_rel_abund <- tsne_otu_names %>% 
  left_join(total_rel_abund, by="asv_id")

@@

mds_asv_plot <- mds_coords %>% 
  left_join(taxonomy, by="asv_id") %>% 
  left_join(asv_otu_key, by="asv_id") %>% 
  left_join(total_rel_abund, by="asv_id") %>% 
  filter(!(asv_id %in% c("639", "534", "646"))) # remove Trebouxiophyceae outliers

@@

Run IQtree
```{r tree, fig.width=8, fig.height=20}
# root tree with Trebouxiophyceae
root <- taxonomy %>% 
  filter(class=="Trebouxiophyceae") %>% 
  head(1) %>% 
  pull(asv_id)

@@

# clades annotated by OTU
otu_tree <- rbcl_ggtree %<+% asv_otu_key + 
  geom_tippoint(aes(color=otu_name, alpha=0.7), size=3) + # 
  geom_text2(data=rbcl_ggtree_data, aes(label=paste0(boot,";",bayes)),  vjust= 0.01, hjust = 1, size=3) + # add bootstrap vals
  theme(legend.position = c(0.2,0.7), 
        legend.title = element_blank(), 
        legend.key = element_blank())

otu_tree

@@

ggsave(filename = here("manuscript/figs/2_taxonomy/tsne.png"), plot=final_tsne, height = 6, width = 8)
ggsave(filename = here("manuscript/figs/2_taxonomy/mds.png"), plot=mds,  height = 6, width = 8)
ggsave(filename = here("manuscript/figs/tree.png"), plot=otu_tree, width=8, height = 20)
```

>>>>

otus_asv_name <- otus %>% 
  left_join(otu_names, by="otu") %>% 
  select(asv_id, otu_name)

tsne_added_names <- tsne_edited_otus %>% 
  left_join(otus_asv_name, by="asv_id")
```

```{r echo=F, fig.width=8, fig.height=6}
ggplot(tsne_added_names, aes(x=V1, y=V2)) +
  ggforce::geom_mark_ellipse(aes(fill=otu_name, label=otu_name), alpha=0.3, show.legend = FALSE) +
  # plot points not assigned to genus or species
  geom_point(data = filter(tsne_added_names, is.na(plot_taxa)), 
             color = "grey",alpha = 0.8, size = 3) +
  # color points that were assigned to genus or species
  geom_point(data = filter(tsne_added_names, !is.na(plot_taxa)), 
             aes(color = plot_taxa %>% fct_relevel(taxa.order)), alpha = 0.8, size = 3) +
  labs(title = "Points colored by both genus and species", color="ASV taxonomic assignment") +
  theme_few() 
```

@@

tsne_rel_abund <- tsne_added_names %>% 
  left_join(total_rel_abund, by="asv_id")

@@

mds_asv_plot <- mds_coords %>% 
  left_join(taxonomy, by="asv_id") %>% 
  left_join(otus_asv_name, by="asv_id") %>% 
  left_join(total_rel_abund, by="asv_id") %>% 
  filter(!(asv_id %in% c("639", "534", "646"))) # remove Trebouxiophyceae outliers

@@

Run IQtree
```{r tree, fig.width=7, fig.height=10}
# root tree with Trebouxiophyceae
root <- taxonomy %>% 
  filter(class=="Trebouxiophyceae") %>% 
  head(1) %>% 
  pull(asv_id)

@@

# clades annotated by OTU
otu_tree <- rbcl_ggtree %<+% otus_asv_name + 
  geom_tippoint(aes(color=otu_name, alpha=0.7), size=3) + # 
  geom_text2(data=rbcl_ggtree_data, aes(label=paste0(boot,";",bayes)),  vjust= 0.01, hjust = 1, size=3) + # add bootstrap vals
  theme(legend.position = c(0.2,0.7), 
        legend.title = element_blank(), 
        legend.key = element_blank())

taxonomy_tree
otu_tree

@@

ggsave(filename = here("manuscript/figs/2_taxonomy/tsne.png"), plot=final_tsne, height = 6, width = 8)
ggsave(filename = here("manuscript/figs/2_taxonomy/mds.png"), plot=mds,  height = 6, width = 8)

ggsave(filename = here("manuscript/figs/otu_tree.png"), plot=otu_tree, width=7, height = 10)
ggsave(filename = here("manuscript/figs/taxonomy_tree.png"), plot=taxonomy_tree, width=7, height = 10) 
```
